---
title: "Who specialises in climate adaptation technology?"
subtitle: "Revealed technological advantage vs. sheer size, for adaptation innovation"
author: "Ralf"
date: "`r format(Sys.Date(), '%B %Y')`"
params:
  top_n: 10
  iseapp_data: "../iseapp/inst/extdata"
output:
  html_document:
    self_contained: true
    theme: default
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center"
)
library(DBI)
library(duckdb)
library(dplyr)
library(ggplot2)
library(ggiraph)
library(countrycode)
```

```{r params}
# ── Inputs ────────────────────────────────────────────────────────────────────
# top_n         : how many countries each panel shows
# min_allinnos  : drop countries with fewer than this many innovations in total
top_n        <- params$top_n
min_allinnos <- 100

# Filters, matching the Innovation Strategy Explorer settings this note is
# built from: multi-application families only, utility models excluded,
# no granted-only restriction, no firm or city filter.
multifam_only <- TRUE
exclude_um    <- TRUE
granted_only  <- FALSE
toflow        <- "is_global"
tech_sel      <- "Adaptation"
```

## The question

Adaptation technology — irrigation and water management, flood and coastal defence,
drought- and heat-resistant agriculture, early-warning systems, cooling, resilient
construction — is the part of climate innovation that matters most to the countries
already living with the damage. So: **which countries are specialised in it?**

There are two very different answers, and the figure below puts them side by side.

- **Revealed technological advantage (RTA)** asks whether adaptation takes up a
  *larger share of a country's own innovation* than it does of world innovation.
  It is a measure of orientation, and it is scale-free — a small country can score high.
- **Share** asks how much of *the world's* adaptation innovation a country accounts for.
  That is a measure of weight, and it is dominated by the big patenting economies.

The countries that top the first list barely register on the second. That gap is
the story.

## How the numbers are built

Everything comes from the [Innovation Strategy Explorer](https://mondpanther-iseapp2.share.connect.posit.cloud/)
patent database (DOCDB families, inventor country). An "innovation" is a patent family
counted against each country whose inventors contributed to it.

RTA is the normalised (Balassa) form the Explorer uses,

$$\text{RTA}_c \;=\; \frac{2\,s_c}{s_c + s}, \qquad
s_c = \frac{\text{adaptation innovations in } c}{\text{all innovations in } c}, \qquad
s = \frac{\text{adaptation innovations, all countries}}{\text{all innovations, all countries}}$$

which sits on $[0, 2]$ with **1 = exactly the world average orientation**.

Three restrictions, all of which matter:

- **Multi-application families only** — the family must have been filed in more than one
  jurisdiction, a standard quality screen that removes purely domestic one-off filings.
- **Utility models excluded** — families where *every* application is a utility model are
  dropped. About 30% of PATSTAT families are flagged this way, heavily concentrated in
  China, and they are much weaker rights than patents.
- **Countries with fewer than `r min_allinnos` innovations in total are excluded**, so the
  RTA ranking is not won by a country with four patents, two of which happen to be about
  irrigation.

```{r data}
d <- normalizePath(params$iseapp_data, mustWork = TRUE)

con <- dbConnect(duckdb())

# `is_global` is derived on the fly exactly as the Explorer's
# full_patent_database view derives it; only its NOT NULL-ness is used here,
# as the row-selection restriction the app applies to every query.
dbExecute(con, sprintf("
  CREATE OR REPLACE VIEW p_raw AS
  SELECT *,
    CASE WHEN cost IS NULL OR cost = 0 THEN 0
         ELSE ((alpha + 1) / cost) * ev_global * CAST(pv <= 2 * cost AS DOUBLE)
    END AS is_global
  FROM read_parquet('%s')", file.path(d, "patent_database.parquet")))

for (t in c("patents_x_tech", "tech_lookup", "country_lookup")) {
  dbExecute(con, sprintf(
    "CREATE OR REPLACE TABLE %s AS SELECT * FROM read_parquet('%s')",
    t, file.path(d, paste0(t, ".parquet"))))
}

flt <- paste(
  if (multifam_only) "AND p.fam_size_min2 = TRUE" else "",
  if (exclude_um)    "AND p.is_um = FALSE"        else "",
  if (granted_only)  "AND p.granted = TRUE"       else "")

# Numerator: distinct families per country matching the technology filter.
tech_counts <- dbGetQuery(con, sprintf("
  WITH ft AS (
    SELECT DISTINCT t.docdb_family_id
    FROM patents_x_tech t
    JOIN tech_lookup tl ON t.technology = tl.technology
    WHERE tl.tech_group IN ('%s') OR t.technology IN ('%s')
  )
  SELECT p.ctry_code, COUNT(DISTINCT p.docdb_family_id) AS innos
  FROM p_raw p
  INNER JOIN ft ON p.docdb_family_id = ft.docdb_family_id
  JOIN country_lookup cl ON p.ctry_code = cl.ctry_code
  WHERE p.%s IS NOT NULL %s
  GROUP BY p.ctry_code", tech_sel, tech_sel, toflow, flt))

# Denominator: the same row base, without the technology filter.
all_counts <- dbGetQuery(con, sprintf("
  SELECT p.ctry_code, COUNT(DISTINCT p.docdb_family_id) AS allinnos
  FROM p_raw p
  JOIN country_lookup cl ON p.ctry_code = cl.ctry_code
  WHERE p.%s IS NOT NULL %s
  GROUP BY p.ctry_code", toflow, flt))

lookup <- dbGetQuery(con, "SELECT ctry_code, is_hic FROM country_lookup")

world_share <- sum(tech_counts$innos) / sum(all_counts$allinnos)

dat <- tech_counts |>
  left_join(all_counts, by = "ctry_code") |>
  left_join(lookup,     by = "ctry_code") |>
  filter(allinnos >= min_allinnos) |>
  mutate(
    share_c      = innos / allinnos,
    RTA          = 2 * share_c / (share_c + world_share),
    share_global = 100 * innos / sum(tech_counts$innos),
    income       = if_else(is_hic, "High income countries",
                                   "Low & middle income countries"),
    country_name = coalesce(
      countrycode(ctry_code, "iso2c", "country.name.en", warn = FALSE),
      ctry_code)
  )
```

The current build gives `r format(sum(tech_counts$innos), big.mark = ",")` country-level
adaptation innovations against
`r format(sum(all_counts$allinnos), big.mark = ",")` innovations in total — so adaptation
is **`r sprintf("%.2f%%", 100 * world_share)`** of world innovation, and that figure is the
yardstick RTA = 1 refers to. After the `r min_allinnos`-innovation floor,
`r nrow(dat)` countries remain.

## The figure

Each bar is one country. **Height is RTA**; **width is the country's share of world
adaptation innovation**, so the bars are laid out along an axis measured in percent
of world adaptation innovation — a bar you can barely see is a country that barely
registers in the global total. Colour is the Explorer's income classification. Hover
any bar for the underlying counts.

```{r disconnect}
dbDisconnect(con, shutdown = TRUE)
```

```{r plotfun}
pal <- c("High income countries"         = "#3d5bd9",
         "Low & middle income countries" = "#c8bb8b")

lab_col <- c("High income countries"         = "#2b3fa8",
             "Low & middle income countries" = "#8a7c46")

# Spread label positions so narrow neighbouring bars stay readable: labels are
# nudged apart left-to-right until they no longer overlap, then pulled back
# inside the panel, and leader lines run from each bar to its label.
spread_labels <- function(x, halfw, lo, hi, pad) {
  o <- order(x); x <- x[o]; halfw <- halfw[o]
  for (i in seq_along(x)[-1]) {
    need <- x[i - 1] + halfw[i - 1] + halfw[i] + pad
    if (x[i] < need) x[i] <- need
  }
  if (x[length(x)] + halfw[length(x)] > hi) {
    x[length(x)] <- hi - halfw[length(x)]
    for (i in rev(seq_along(x))[-1]) {
      room <- x[i + 1] - halfw[i + 1] - halfw[i] - pad
      if (x[i] > room) x[i] <- room
    }
  }
  x <- pmax(x, lo + halfw)
  out <- numeric(length(x)); out[o] <- x; out
}

# Variable-width bar panel: width proportional to `share_global`, laid out
# cumulatively along x, height = RTA, labels on leader lines above the bars.
make_panel <- function(df, title) {
  df <- df |>
    arrange(desc(RTA)) |>
    mutate(
      xmax = cumsum(share_global),
      xmin = xmax - share_global,
      xmid = (xmin + xmax) / 2
    )

  span <- max(df$xmax)

  # Three staggered rows above the bars, so a label never sits on a bar.
  rows  <- c(2.12, 2.30, 2.48)
  df$lab_y <- rows[(rank(df$xmid, ties.method = "first") - 1) %% 3 + 1]

  # Rough character width in data units: the drawing area holds roughly 60
  # characters of label text across at this font size once the legend is
  # accounted for.
  char_w   <- span / 60
  df$halfw <- nchar(df$country_name) * char_w / 2
  df$lab_x <- df$xmid

  for (r in rows) {
    k <- df$lab_y == r
    if (sum(k) > 1) {
      df$lab_x[k] <- spread_labels(df$xmid[k], df$halfw[k], 0, span, char_w * 0.4)
    } else if (sum(k) == 1) {
      df$lab_x[k] <- pmin(pmax(df$xmid[k], df$halfw[k]), span - df$halfw[k])
    }
  }

  ggplot(df) +
    geom_rect_interactive(
      aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = RTA, fill = income,
          data_id = ctry_code,
          tooltip = paste0(
            country_name,
            "\nRTA: ", sprintf("%.2f", RTA),
            "\nAdaptation innovations: ", format(innos, big.mark = ","),
            "\nAll innovations: ", format(allinnos, big.mark = ","),
            "\nShare of world adaptation: ", sprintf("%.2f%%", share_global),
            "\nAdaptation share of own innovation: ", sprintf("%.2f%%", 100 * share_c))),
      colour = "white", linewidth = 0.15) +
    # leader line: up from the bar, then across to the label
    geom_segment(aes(x = xmid, xend = xmid, y = RTA, yend = 2.04),
                 colour = "#9a9ab0", linewidth = 0.3) +
    geom_segment(aes(x = xmid, xend = lab_x, y = 2.04, yend = lab_y - 0.04),
                 colour = "#9a9ab0", linewidth = 0.3) +
    geom_text(aes(x = lab_x, y = lab_y, label = country_name, colour = income),
              size = 3.1, vjust = 0, show.legend = FALSE) +
    geom_hline(yintercept = 1, linetype = "dashed",
               colour = "#8a8a8a", linewidth = 0.35) +
    scale_fill_manual(values = pal, name = NULL, drop = FALSE) +
    scale_colour_manual(values = lab_col, drop = FALSE) +
    scale_y_continuous(breaks = c(0, 1, 2)) +
    coord_cartesian(xlim = c(0, span), ylim = c(0, 2.62),
                    expand = FALSE, clip = "off") +
    labs(title = title,
         x = "Cumulative share of world adaptation innovation (%)",
         y = "RTA") +
    theme_minimal(base_size = 12) +
    theme(
      plot.title         = element_text(hjust = 0.5, face = "bold", size = 13),
      plot.margin        = margin(6, 10, 6, 6),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.title.x       = element_text(size = 10, colour = "#555"),
      axis.title.y       = element_text(size = 10, colour = "#555", angle = 0,
                                        vjust = 0.53, hjust = 1,
                                        margin = margin(r = 4)),
      legend.position    = "right",
      legend.text        = element_text(size = 9)
    )
}

top_rta   <- dat |> arrange(desc(RTA))          |> head(top_n)
top_share <- dat |> arrange(desc(share_global)) |> head(top_n)
```

```{r fig_rta, fig.width = 9, fig.height = 4.6}
girafe(ggobj = make_panel(top_rta,
         sprintf("Top %d countries by RTA in adaptation", top_n)),
       width_svg = 9, height_svg = 4.6,
       options = list(
         opts_sizing(rescale = TRUE, width = 1),
         opts_hover(css = "fill-opacity:0.75;"),
         opts_tooltip(css = paste0("background-color:white;padding:6px;",
                                   "border-radius:3px;border:1px solid #ccc;",
                                   "font-family:sans-serif;font-size:11px;"))))
```

```{r fig_share, fig.width = 9, fig.height = 4.6}
girafe(ggobj = make_panel(top_share,
         sprintf("Top %d countries by share of world adaptation innovation", top_n)),
       width_svg = 9, height_svg = 4.6,
       options = list(
         opts_sizing(rescale = TRUE, width = 1),
         opts_hover(css = "fill-opacity:0.75;"),
         opts_tooltip(css = paste0("background-color:white;padding:6px;",
                                   "border-radius:3px;border:1px solid #ccc;",
                                   "font-family:sans-serif;font-size:11px;"))))
```

Note the two x-axes. The top panel's `r top_n` countries together account for
**`r sprintf("%.2f%%", sum(top_rta$share_global))`** of world adaptation innovation;
the bottom panel's `r top_n` account for
**`r sprintf("%.1f%%", sum(top_share$share_global))`**.

## What it says

**Specialisation and scale are almost disjoint.** The countries most *oriented* towards
adaptation are small and mostly outside the high-income group; between them they hold a
rounding error's worth of the world's adaptation innovation. The countries that *hold*
the adaptation portfolio are the usual patent powers, and their RTAs cluster around 1 —
adaptation is neither a priority nor a blind spot for them, it is simply proportional
to everything else they do.

**Around 1 is the norm, and that is itself informative.** Nobody in the bottom panel is
strongly specialised. There is no large economy that has turned itself into an adaptation
innovator the way some have in, say, green energy.

**The demand-supply mismatch is visible in one picture.** Adaptation need is concentrated
where climate exposure and low incomes coincide. Adaptation *innovation capacity* is
concentrated where the patent system is. The countries in the top panel are pointed in
the right direction but are tiny; the countries in the bottom panel are large but not
pointed anywhere in particular.

## The numbers

### Top `r top_n` by RTA

```{r tbl_rta}
top_rta |>
  transmute(Country = country_name,
            Income  = income,
            RTA     = round(RTA, 2),
            `Adaptation innovations` = innos,
            `All innovations`        = allinnos,
            `Adaptation share of own innovation (%)` = round(100 * share_c, 2),
            `Share of world adaptation (%)`          = round(share_global, 3))
```

### Top `r top_n` by share of world adaptation innovation

```{r tbl_share}
top_share |>
  transmute(Country = country_name,
            Income  = income,
            RTA     = round(RTA, 2),
            `Adaptation innovations` = innos,
            `All innovations`        = allinnos,
            `Adaptation share of own innovation (%)` = round(100 * share_c, 2),
            `Share of world adaptation (%)`          = round(share_global, 3))
```

## Caveats

**Income classification is binary here.** The Explorer's database carries one income
flag per country (`is_hic` in `country_lookup`), so the legend splits high income from
low-and-middle income rather than the three World Bank classes. It also follows the
project's own fixed country list, which is not re-based on the World Bank's annual
reclassifications — a handful of countries (Chile among them) sit in the low-and-middle
group here although the World Bank now calls them high income.

**Innovations are counted per country, not split between them.** A family with inventors
in two countries counts once for each, so country counts sum to more than the world
total and the share axis is a share of that summed total.

**"Adaptation" is a CPC-based technology class**, and the boundary between adaptation and
mitigation is genuinely fuzzy — efficient cooling, water treatment and resilient
agriculture can sit on either side.

**The multi-application and utility-model screens are doing real work.** Dropping them
raises China's counts sharply. They are defensible quality screens, not neutral ones.

## Reproducing this in the Explorer

The same cut is reachable interactively:
[Country Explorer → Revealed Technological Advantage, Adaptation, multi-application
families only, utility models excluded, all-innovation threshold 100](https://mondpanther-iseapp2.share.connect.posit.cloud/?_inputs_&country-sidebar=true&navbar_page=%22Country%20Explorer%22&country-inner_tabs=%22Revealed%20Technological%20Advantage%22&country-country=%22All%20countries%22&country-techs=%22Adaptation%22&country-toflow=%22is_global%22&country-multifam_only=true&country-exclude_um=true&country-granted_only=false&country-topn_rta=10&country-minallinnos_rta=100).

This note renders from `adaptation-rta.Rmd` with `top_n` as a parameter:

```r
rmarkdown::render("adaptation-rta.Rmd", params = list(top_n = 15))
```
