---
title: "Who specialises in climate adaptation technology?"
subtitle: "Revealed technological advantage vs. sheer size, for adaptation innovation"
author: "Ralf"
date: "`r format(Sys.Date(), '%B %Y')`"
params:
  top_n: 10
  iseapp_data: "../iseapp/inst/extdata"
output:
  html_document:
    self_contained: true
    theme: default
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center"
)
library(DBI)
library(duckdb)
library(dplyr)
library(jsonlite)
library(countrycode)
```

```{r params}
# ── Inputs ────────────────────────────────────────────────────────────────────
# top_n         : how many countries each panel starts out showing. In the
#                 rendered page this is a slider, so `top_n` only sets where
#                 the slider sits when the page loads.
# min_allinnos  : drop countries with fewer than this many innovations in total
top_n        <- params$top_n
min_allinnos <- 100

# Filters, matching the Innovation Strategy Explorer settings this note is
# built from: multi-application families only, utility models excluded,
# no granted-only restriction, no firm or city filter.
multifam_only <- TRUE
exclude_um    <- TRUE
granted_only  <- FALSE
toflow        <- "is_global"
tech_sel      <- "Adaptation"
```

## The question

Adaptation technology — irrigation and water management, flood and coastal defence,
drought- and heat-resistant agriculture, early-warning systems, cooling, resilient
construction — is the part of climate innovation that matters most to the countries
already living with the damage. So: **which countries are specialised in it?**

There are two very different answers, and the figure below puts them side by side.

- **Revealed technological advantage (RTA)** asks whether adaptation takes up a
  *larger share of a country's own innovation* than it does of world innovation.
  It is a measure of orientation, and it is scale-free — a small country can score high.
- **Share** asks how much of *the world's* adaptation innovation a country accounts for.
  That is a measure of weight, and it is dominated by the big patenting economies.

The countries that top the first list barely register on the second. That gap is
the story.

## How the numbers are built

Everything comes from the [Innovation Strategy Explorer](https://mondpanther-iseapp2.share.connect.posit.cloud/)
patent database (DOCDB families, inventor country). An "innovation" is a patent family
counted against each country whose inventors contributed to it.

RTA is the normalised (Balassa) form the Explorer uses,

$$\text{RTA}_c \;=\; \frac{2\,s_c}{s_c + s}, \qquad
s_c = \frac{\text{adaptation innovations in } c}{\text{all innovations in } c}, \qquad
s = \frac{\text{adaptation innovations, all countries}}{\text{all innovations, all countries}}$$

which sits on $[0, 2]$ with **1 = exactly the world average orientation**.

Three restrictions, all of which matter:

- **Multi-application families only** — the family must have been filed in more than one
  jurisdiction, a standard quality screen that removes purely domestic one-off filings.
- **Utility models excluded** — families where *every* application is a utility model are
  dropped. About 30% of PATSTAT families are flagged this way, heavily concentrated in
  China, and they are much weaker rights than patents.
- **Countries with fewer than `r min_allinnos` innovations in total are excluded**, so the
  RTA ranking is not won by a country with four patents, two of which happen to be about
  irrigation.

```{r data}
d <- normalizePath(params$iseapp_data, mustWork = TRUE)

con <- dbConnect(duckdb())

# `is_global` is derived on the fly exactly as the Explorer's
# full_patent_database view derives it; only its NOT NULL-ness is used here,
# as the row-selection restriction the app applies to every query.
dbExecute(con, sprintf("
  CREATE OR REPLACE VIEW p_raw AS
  SELECT *,
    CASE WHEN cost IS NULL OR cost = 0 THEN 0
         ELSE ((alpha + 1) / cost) * ev_global * CAST(pv <= 2 * cost AS DOUBLE)
    END AS is_global
  FROM read_parquet('%s')", file.path(d, "patent_database.parquet")))

for (t in c("patents_x_tech", "tech_lookup", "country_lookup")) {
  dbExecute(con, sprintf(
    "CREATE OR REPLACE TABLE %s AS SELECT * FROM read_parquet('%s')",
    t, file.path(d, paste0(t, ".parquet"))))
}

flt <- paste(
  if (multifam_only) "AND p.fam_size_min2 = TRUE" else "",
  if (exclude_um)    "AND p.is_um = FALSE"        else "",
  if (granted_only)  "AND p.granted = TRUE"       else "")

# Numerator: distinct families per country matching the technology filter.
tech_counts <- dbGetQuery(con, sprintf("
  WITH ft AS (
    SELECT DISTINCT t.docdb_family_id
    FROM patents_x_tech t
    JOIN tech_lookup tl ON t.technology = tl.technology
    WHERE tl.tech_group IN ('%s') OR t.technology IN ('%s')
  )
  SELECT p.ctry_code, COUNT(DISTINCT p.docdb_family_id) AS innos
  FROM p_raw p
  INNER JOIN ft ON p.docdb_family_id = ft.docdb_family_id
  JOIN country_lookup cl ON p.ctry_code = cl.ctry_code
  WHERE p.%s IS NOT NULL %s
  GROUP BY p.ctry_code", tech_sel, tech_sel, toflow, flt))

# Denominator: the same row base, without the technology filter.
all_counts <- dbGetQuery(con, sprintf("
  SELECT p.ctry_code, COUNT(DISTINCT p.docdb_family_id) AS allinnos
  FROM p_raw p
  JOIN country_lookup cl ON p.ctry_code = cl.ctry_code
  WHERE p.%s IS NOT NULL %s
  GROUP BY p.ctry_code", toflow, flt))

lookup <- dbGetQuery(con, "SELECT ctry_code, is_hic FROM country_lookup")

world_share <- sum(tech_counts$innos) / sum(all_counts$allinnos)

dat <- tech_counts |>
  left_join(all_counts, by = "ctry_code") |>
  left_join(lookup,     by = "ctry_code") |>
  filter(allinnos >= min_allinnos) |>
  mutate(
    share_c      = innos / allinnos,
    RTA          = 2 * share_c / (share_c + world_share),
    share_global = 100 * innos / sum(tech_counts$innos),
    income       = if_else(is_hic, "High income countries",
                                   "Low & middle income countries"),
    country_name = coalesce(
      countrycode(ctry_code, "iso2c", "country.name.en", warn = FALSE),
      ctry_code)
  )
```

The current build gives `r format(sum(tech_counts$innos), big.mark = ",")` country-level
adaptation innovations against
`r format(sum(all_counts$allinnos), big.mark = ",")` innovations in total — so adaptation
is **`r sprintf("%.2f%%", 100 * world_share)`** of world innovation, and that figure is the
yardstick RTA = 1 refers to. After the `r min_allinnos`-innovation floor,
`r nrow(dat)` countries remain.

## The figure

Each bar is one country. **Height is RTA**; **width is the country's share of world
adaptation innovation**, so the bars are laid out along an axis measured in percent
of world adaptation innovation — a bar you can barely see is a country that barely
registers in the global total. Colour is the Explorer's income classification. Hover
any bar for the underlying counts, and **move the slider** to change how many
countries each panel shows.

```{r disconnect}
dbDisconnect(con, shutdown = TRUE)
```

```{r widget_assets, include = FALSE}
# Markup, styling and the drawing code for the interactive figure. Kept out of
# the output chunk so the piece that actually emits anything stays short.
# NOTE: no blank lines in the emitted HTML — pandoc would otherwise break out
# of the raw-HTML block and start interpreting the contents as markdown.
adapt_widget_html_head <- paste0(
'<style>',
'.adapt{font-family:-apple-system,BlinkMacSystemFont,"Segoe UI",Helvetica,Arial,sans-serif;margin:16px 0 4px;position:relative;}',
'.adapt-bar{display:flex;flex-wrap:wrap;align-items:center;gap:18px;padding:10px 14px;border:1px solid #e3e3df;border-radius:5px;background:#faf9f6;font-size:13px;color:#333;}',
'.adapt-bar label{font-weight:600;}',
'.adapt-bar input[type=range]{vertical-align:middle;width:190px;}',
'.adapt-n{font-variant-numeric:tabular-nums;font-weight:600;min-width:1.4em;display:inline-block;}',
'.adapt-key{display:inline-flex;align-items:center;gap:6px;}',
'.adapt-sw{width:13px;height:13px;border-radius:2px;display:inline-block;}',
'.adapt-panel{margin-top:10px;}',
'.adapt-tip{position:absolute;pointer-events:none;opacity:0;transition:opacity .08s;background:#fff;border:1px solid #ccc;border-radius:3px;padding:6px 8px;font-size:11px;line-height:1.45;white-space:pre;box-shadow:0 1px 4px rgba(0,0,0,.14);z-index:20;}',
'.adapt-btn{font:inherit;font-size:12px;padding:5px 11px;border:1px solid #c9c9c4;border-radius:4px;background:#fff;color:#2b5a8c;cursor:pointer;}',
'.adapt-btn:hover{background:#f0f4f9;border-color:#2b5a8c;}',
'.adapt-cap{font-size:13px;color:#555;margin:14px 0 6px;display:flex;flex-wrap:wrap;gap:12px;align-items:center;justify-content:space-between;}',
'.adapt-table{width:100%;border-collapse:collapse;font-size:12.5px;}',
'.adapt-table th{text-align:right;border-bottom:2px solid #d8d8d3;padding:5px 7px;font-weight:600;}',
'.adapt-table td{text-align:right;border-bottom:1px solid #eeeee9;padding:4px 7px;font-variant-numeric:tabular-nums;}',
'.adapt-table th:first-child,.adapt-table td:first-child,',
'.adapt-table th:nth-child(2),.adapt-table td:nth-child(2){text-align:left;}',
'</style>',
'<div class="adapt" id="adapt-app">',
'<div class="adapt-bar">',
'<span><label for="adapt-slider">Countries shown per panel:</label> ',
'<span class="adapt-n" id="adapt-n-out"></span></span>',
'<input type="range" id="adapt-slider" min="3" max="40" step="1">',
'<span class="adapt-key"><span class="adapt-sw" style="background:#3d5bd9"></span>High income countries</span>',
'<span class="adapt-key"><span class="adapt-sw" style="background:#c8bb8b"></span>Low &amp; middle income countries</span>',
'<button class="adapt-btn" id="adapt-dl-all" type="button">Download all countries (CSV)</button>',
'</div>',
'<div class="adapt-panel" id="adapt-chart-rta"></div>',
'<div class="adapt-panel" id="adapt-chart-share"></div>',
'<div class="adapt-tip" id="adapt-tip"></div>',
'</div>')

adapt_widget_js <- r"(<script>
(function(){
var NS="http://www.w3.org/2000/svg";
var PAL={"High income countries":"#3d5bd9","Low & middle income countries":"#c8bb8b"};
var LABCOL={"High income countries":"#2b3fa8","Low & middle income countries":"#8a7c46"};
var COLS=[["country","Country",null],["income","Income",null],["rta","RTA",2],
          ["innos","Adaptation innovations",0],["allinnos","All innovations",0],
          ["share_own","Adaptation share of own innovation (%)",2],
          ["share_world","Share of world adaptation (%)",3]];
var ROWS=ADAPT_CFG.rows, cur={};
function mk(t,a){var e=document.createElementNS(NS,t);for(var k in a){e.setAttribute(k,a[k]);}return e;}
function num(x,d){return d===0?Math.round(x).toLocaleString("en-US"):x.toFixed(d);}
// top n on the chosen criterion, then laid out left-to-right by descending RTA
function pick(mode,n){
  var s=ROWS.slice();
  s.sort(function(a,b){return mode==="rta"?b.rta-a.rta:b.share_world-a.share_world;});
  s=s.slice(0,n).map(function(r){var o={};for(var k in r){o[k]=r[k];}return o;});
  s.sort(function(a,b){return b.rta-a.rta;});
  var x=0;
  s.forEach(function(r){r.x0=x;x+=r.share_world;r.x1=x;r.xm=(r.x0+r.x1)/2;});
  return s;
}
function ticks(span){
  var raw=span/5,mag=Math.pow(10,Math.floor(Math.log(raw)/Math.LN10)),nm=raw/mag,st;
  if(nm<1.5){st=mag;}else if(nm<3){st=2*mag;}else if(nm<7){st=5*mag;}else{st=10*mag;}
  var out=[],v=0,guard=0;
  while(v<=span*1.0000001&&guard++<60){out.push(v);v+=st;}
  return out;
}
// nudge labels apart in pixel space, then keep them inside the drawing area
function spread(items,lo,hi,pad){
  items.sort(function(a,b){return a.x-b.x;});
  for(var i=1;i<items.length;i++){
    var need=items[i-1].x+items[i-1].w/2+items[i].w/2+pad;
    if(items[i].x<need){items[i].x=need;}
  }
  var last=items[items.length-1];
  if(last&&last.x+last.w/2>hi){
    last.x=hi-last.w/2;
    for(var j=items.length-2;j>=0;j--){
      var room=items[j+1].x-items[j+1].w/2-items[j].w/2-pad;
      if(items[j].x>room){items[j].x=room;}
    }
  }
  items.forEach(function(it){if(it.x-it.w/2<lo){it.x=lo+it.w/2;}});
}
function tipText(r){
  return r.country+"\nRTA: "+num(r.rta,2)
    +"\nAdaptation innovations: "+num(r.innos,0)
    +"\nAll innovations: "+num(r.allinnos,0)
    +"\nShare of world adaptation: "+num(r.share_world,2)+"%"
    +"\nAdaptation share of own innovation: "+num(r.share_own,2)+"%";
}
function drawPanel(host,rows,title){
  host.innerHTML="";
  var app=document.getElementById("adapt-app"),tip=document.getElementById("adapt-tip");
  var W=Math.max(host.clientWidth||680,300),H=330,m={l:46,r:16,t:88,b:48};
  var svg=mk("svg",{width:"100%",height:H,viewBox:"0 0 "+W+" "+H});
  host.appendChild(svg);
  var span=rows.length?rows[rows.length-1].x1:1;
  var px=function(v){return m.l+(v/span)*(W-m.l-m.r);};
  var py=function(v){return (H-m.b)-(v/2.05)*(H-m.b-m.t);};
  function txt(x,y,s,o){
    var t=mk("text",{x:x,y:y,"font-size":(o&&o.size)||11,fill:(o&&o.fill)||"#555",
      "text-anchor":(o&&o.anchor)||"middle"});
    if(o&&o.weight){t.setAttribute("font-weight",o.weight);}
    t.textContent=s;svg.appendChild(t);return t;
  }
  txt(W/2,20,title,{size:13,fill:"#1d1d1d",weight:"600"});
  [0,1,2].forEach(function(v){
    svg.appendChild(mk("line",{x1:m.l,x2:W-m.r,y1:py(v),y2:py(v),
      stroke:v===1?"#8a8a8a":"#e6e6e2","stroke-width":v===1?1:1,
      "stroke-dasharray":v===1?"4 3":"none"}));
    txt(m.l-7,py(v)+4,String(v),{anchor:"end"});
  });
  txt(m.l-7,py(2)-12,"RTA",{anchor:"end",size:10});
  svg.appendChild(mk("line",{x1:m.l,x2:W-m.r,y1:py(0),y2:py(0),stroke:"#bdbdb8"}));
  ticks(span).forEach(function(v){
    txt(px(v),py(0)+16,span<3?v.toFixed(2):(span<30?v.toFixed(1):String(Math.round(v))));
  });
  txt((m.l+W-m.r)/2,H-10,"Cumulative share of world adaptation innovation (%)",{size:10});
  rows.forEach(function(r){
    var x=px(r.x0),w=Math.max(px(r.x1)-px(r.x0),0.8);
    var rect=mk("rect",{x:x,y:py(r.rta),width:w,height:py(0)-py(r.rta),
      fill:PAL[r.income]||"#999",stroke:"#fff","stroke-width":0.6});
    rect.style.cursor="crosshair";
    rect.addEventListener("mouseenter",function(){rect.setAttribute("fill-opacity",0.72);});
    rect.addEventListener("mousemove",function(ev){
      var b=app.getBoundingClientRect();
      tip.textContent=tipText(r);tip.style.opacity=1;
      var tx=ev.clientX-b.left+14,ty=ev.clientY-b.top+12;
      tip.style.left=Math.min(tx,b.width-tip.offsetWidth-6)+"px";tip.style.top=ty+"px";
    });
    rect.addEventListener("mouseleave",function(){
      rect.setAttribute("fill-opacity",1);tip.style.opacity=0;});
    svg.appendChild(rect);
  });
  // labels: three staggered rows above the bars, joined by leader lines
  var bands=[m.t-14,m.t-32,m.t-50],slots=[[],[],[]];
  rows.forEach(function(r,i){
    var b=i%3,t=txt(0,bands[b],r.country,{fill:LABCOL[r.income]||"#555",size:10.5});
    var w=0;try{w=t.getComputedTextLength();}catch(e){w=r.country.length*5.6;}
    if(!w){w=r.country.length*5.6;}
    slots[b].push({x:px(r.xm),w:w,node:t,bar:px(r.xm),y:bands[b],rta:r.rta});
  });
  slots.forEach(function(items){spread(items,m.l,W-m.r,7);});
  slots.forEach(function(items){items.forEach(function(it){
    it.node.setAttribute("x",it.x);
    svg.insertBefore(mk("line",{x1:it.bar,x2:it.bar,y1:py(it.rta),y2:m.t-8,
      stroke:"#9a9ab0","stroke-width":0.7}),svg.firstChild);
    svg.insertBefore(mk("line",{x1:it.bar,x2:it.x,y1:m.t-8,y2:it.y+3,
      stroke:"#9a9ab0","stroke-width":0.7}),svg.firstChild);
  });});
}
function tableHTML(rows){
  var h="<table class='adapt-table'><thead><tr>";
  COLS.forEach(function(c){h+="<th>"+c[1]+"</th>";});
  h+="</tr></thead><tbody>";
  rows.forEach(function(r){
    h+="<tr>";
    COLS.forEach(function(c){
      h+="<td>"+(c[2]===null?r[c[0]]:num(r[c[0]],c[2]))+"</td>";
    });
    h+="</tr>";
  });
  return h+"</tbody></table>";
}
function csv(rows){
  var q=function(v){v=String(v);return /[",\n]/.test(v)?'"'+v.replace(/"/g,'""')+'"':v;};
  var out=[COLS.map(function(c){return q(c[1]);}).join(",")];
  rows.forEach(function(r){
    out.push(COLS.map(function(c){
      return q(c[2]===null?r[c[0]]:(c[2]===0?r[c[0]]:r[c[0]].toFixed(c[2])));
    }).join(","));
  });
  return out.join("\n");
}
function download(name,text){
  var b=new Blob(["﻿"+text],{type:"text/csv;charset=utf-8;"});
  var u=URL.createObjectURL(b),a=document.createElement("a");
  a.href=u;a.download=name;document.body.appendChild(a);a.click();
  document.body.removeChild(a);setTimeout(function(){URL.revokeObjectURL(u);},800);
}
function sumShare(rows){
  return rows.reduce(function(s,r){return s+r.share_world;},0);
}
function render(){
  var n=+document.getElementById("adapt-slider").value;
  document.getElementById("adapt-n-out").textContent=n;
  cur.rta=pick("rta",n);cur.share=pick("share",n);
  drawPanel(document.getElementById("adapt-chart-rta"),cur.rta,
            "Top "+n+" countries by RTA in adaptation");
  drawPanel(document.getElementById("adapt-chart-share"),cur.share,
            "Top "+n+" countries by share of world adaptation innovation");
  var t1=document.getElementById("adapt-tbl-rta"),t2=document.getElementById("adapt-tbl-share");
  if(t1){t1.innerHTML=tableHTML(cur.rta);}
  if(t2){t2.innerHTML=tableHTML(cur.share);}
  Array.prototype.forEach.call(document.querySelectorAll(".adapt-nref"),function(e){
    e.textContent=n;});
  var s1=document.getElementById("adapt-sum-rta"),s2=document.getElementById("adapt-sum-share");
  if(s1){s1.textContent=sumShare(cur.rta).toFixed(2)+"%";}
  if(s2){s2.textContent=sumShare(cur.share).toFixed(1)+"%";}
}
function wire(id,fn){var e=document.getElementById(id);if(e){e.addEventListener("click",fn);}}
function init(){
  var sl=document.getElementById("adapt-slider");
  sl.max=ADAPT_CFG.nmax;sl.value=Math.min(ADAPT_CFG.n0,ADAPT_CFG.nmax);
  sl.addEventListener("input",render);
  wire("adapt-dl-all",function(){download("adaptation_rta_all_countries.csv",csv(ROWS));});
  wire("adapt-dl-rta",function(){
    download("adaptation_top"+cur.rta.length+"_by_rta.csv",csv(cur.rta));});
  wire("adapt-dl-share",function(){
    download("adaptation_top"+cur.share.length+"_by_share.csv",csv(cur.share));});
  render();
  var t;window.addEventListener("resize",function(){
    clearTimeout(t);t=setTimeout(render,150);});
}
if(document.readyState==="loading"){document.addEventListener("DOMContentLoaded",init);}
else{init();}
})();
</script>)"
```

```{r widget, results = "asis"}
# ── Interactive figure ───────────────────────────────────────────────────────
# The rendered page is a static, self-contained HTML file — there is no Shiny
# server behind it — so the full country table is embedded as JSON and the two
# panels, the tables and the CSV exports are all drawn client-side from it.
# That is what lets `n` be a live slider rather than a render-time constant.
payload <- dat |>
  arrange(desc(RTA)) |>
  transmute(code        = ctry_code,
            country     = country_name,
            income      = income,
            rta         = round(RTA, 6),
            innos       = as.integer(innos),
            allinnos    = as.integer(allinnos),
            share_own   = round(100 * share_c, 6),
            share_world = round(share_global, 8))

cfg <- list(rows = payload, n0 = top_n, nmax = min(40L, nrow(payload)),
            minAllinnos = min_allinnos)

cat(paste0(
  adapt_widget_html_head,
  "<script>var ADAPT_CFG = ",
  toJSON(cfg, dataframe = "rows", auto_unbox = TRUE, digits = NA),
  ";</script>",
  adapt_widget_js))
```

Note the two x-axes. The top panel's <span class="adapt-nref"></span> countries together
account for <b><span id="adapt-sum-rta"></span></b> of world adaptation innovation; the
bottom panel's <span class="adapt-nref"></span> account for
<b><span id="adapt-sum-share"></span></b>.

## What it says

**Specialisation and scale are almost disjoint.** The countries most *oriented* towards
adaptation are small and mostly outside the high-income group; between them they hold a
rounding error's worth of the world's adaptation innovation. The countries that *hold*
the adaptation portfolio are the usual patent powers, and their RTAs cluster around 1 —
adaptation is neither a priority nor a blind spot for them, it is simply proportional
to everything else they do.

**Around 1 is the norm, and that is itself informative.** Nobody in the bottom panel is
strongly specialised. There is no large economy that has turned itself into an adaptation
innovator the way some have in, say, green energy.

**The demand-supply mismatch is visible in one picture.** Adaptation need is concentrated
where climate exposure and low incomes coincide. Adaptation *innovation capacity* is
concentrated where the patent system is. The countries in the top panel are pointed in
the right direction but are tiny; the countries in the bottom panel are large but not
pointed anywhere in particular.

## The numbers

### By RTA

```{r tbl_rta, results = "asis"}
cat(paste0(
  '<div class="adapt-cap"><span>Top <span class="adapt-nref"></span> countries by RTA',
  ' in adaptation</span>',
  '<button class="adapt-btn" id="adapt-dl-rta" type="button">Download this table (CSV)',
  '</button></div><div id="adapt-tbl-rta"></div>'))
```

### By share of world adaptation innovation

```{r tbl_share, results = "asis"}
cat(paste0(
  '<div class="adapt-cap"><span>Top <span class="adapt-nref"></span> countries by share',
  ' of world adaptation innovation</span>',
  '<button class="adapt-btn" id="adapt-dl-share" type="button">Download this table (CSV)',
  '</button></div><div id="adapt-tbl-share"></div>'))
```

Both tables follow the slider, and the CSV they export is exactly what is on screen.
The button in the control bar above the figure exports all `r nrow(dat)` countries that
clear the `r min_allinnos`-innovation floor, ranked by RTA.

## Caveats

**Income classification is binary here.** The Explorer's database carries one income
flag per country (`is_hic` in `country_lookup`), so the legend splits high income from
low-and-middle income rather than the three World Bank classes. It also follows the
project's own fixed country list, which is not re-based on the World Bank's annual
reclassifications — a handful of countries (Chile among them) sit in the low-and-middle
group here although the World Bank now calls them high income.

**Innovations are counted per country, not split between them.** A family with inventors
in two countries counts once for each, so country counts sum to more than the world
total and the share axis is a share of that summed total.

**"Adaptation" is a CPC-based technology class**, and the boundary between adaptation and
mitigation is genuinely fuzzy — efficient cooling, water treatment and resilient
agriculture can sit on either side.

**The multi-application and utility-model screens are doing real work.** Dropping them
raises China's counts sharply. They are defensible quality screens, not neutral ones.

## Reproducing this in the Explorer

The same cut is reachable interactively:
[Country Explorer → Revealed Technological Advantage, Adaptation, multi-application
families only, utility models excluded, all-innovation threshold 100](https://mondpanther-iseapp2.share.connect.posit.cloud/?_inputs_&country-sidebar=true&navbar_page=%22Country%20Explorer%22&country-inner_tabs=%22Revealed%20Technological%20Advantage%22&country-country=%22All%20countries%22&country-techs=%22Adaptation%22&country-toflow=%22is_global%22&country-multifam_only=true&country-exclude_um=true&country-granted_only=false&country-topn_rta=10&country-minallinnos_rta=100).

This note renders from `adaptation-rta.Rmd`. The slider does the work in the rendered
page, so `top_n` only sets its starting position:

```r
rmarkdown::render("adaptation-rta.Rmd", params = list(top_n = 15))
```
