---
title: "Who sells cars in Europe? Brand origins since 1970"
subtitle: "European vs Japanese vs Korean vs Chinese brands"
author: "Ralf"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    self_contained: true
    theme: default
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 5.2, dpi = 150
)
library(tidyverse)
library(plotly)
```

## The question

Is there a clean breakdown of European car sales by the *nationality of the brand* —
European, Japanese, Korean and Chinese — going back to 1970, so we can put the
Japanese market entry of the 1970s next to the Chinese one of the 2020s?

Short answer: not as a single official series. You have to assemble it, and you have
to be careful about two very different lenses:

- **Brand nationality** (JATO Dynamics): a car counts as "Japanese" if it wears a
  Japanese badge, regardless of where it was built. This is the lens most people mean.
- **Country of manufacture** (ACEA import statistics): a China-built Tesla or Dacia
  counts as a "Chinese" import. On this basis the Asian shares look *much smaller*,
  because Japanese and Korean makers build heavily inside Europe.

Everything below uses the **brand-nationality** lens for the long-run picture, and
notes the manufacturing lens where it matters. Figures are approximate and
interpolated between documented anchor points — treat the shape of the curves, not
the second decimal, as the message. Pre-2000 numbers are the patchiest.

## Assembled series

```{r data}
# Approximate passenger-car market share (%), Western / EU Europe, by brand origin.
# Anchor points from the sources listed at the foot of this note; intervening years
# are linearly interpolated. Shares are illustrative, not official.

anchors <- tribble(
  ~year, ~Japanese, ~Korean, ~Chinese, ~`US/Other`,
  1970,   1.5,       0.0,     0.0,      13.0,
  1975,   5.0,       0.0,     0.0,      12.5,
  1980,   9.0,       0.0,     0.0,      12.0,
  1985,  11.0,       0.0,     0.0,      12.0,
  1990,  12.0,       0.3,     0.0,      11.0,
  1995,  11.5,       1.5,     0.0,      10.0,
  2000,  12.0,       3.0,     0.0,       9.0,
  2005,  13.1,       4.0,     0.0,       8.5,
  2010,  12.6,       5.0,     0.0,       8.0,
  2015,  12.6,       6.2,     0.1,       7.5,
  2020,  12.0,       7.4,     0.5,       6.5,
  2022,  12.2,       8.4,     1.5,       6.0,
  2024,  12.2,       8.6,     3.1,       6.0,
  2025,  12.0,       8.0,     6.1,       6.0
)

# Interpolate to a yearly grid
grid <- tibble(year = seq(min(anchors$year), max(anchors$year)))
interp <- function(col) approx(anchors$year, anchors[[col]], grid$year)$y

df <- grid %>%
  mutate(
    Japanese  = interp("Japanese"),
    Korean    = interp("Korean"),
    Chinese   = interp("Chinese"),
    `US/Other`= interp("US/Other"),
    European  = 100 - Japanese - Korean - Chinese - `US/Other`
  )

long <- df %>%
  pivot_longer(-year, names_to = "origin", values_to = "share") %>%
  mutate(origin = factor(origin,
           levels = c("European", "Japanese", "Korean", "Chinese", "US/Other")))

pal <- c(European = "#2b5a8c", Japanese = "#c0392b", Korean = "#27ae60",
         Chinese  = "#e0a800", `US/Other` = "#8e8e8e")
```

### Composition over time

```{r stacked}
p_area <- long %>%
  ggplot(aes(year, share, fill = origin,
             text = paste0(origin, ": ", round(share, 1), "%  (", year, ")"))) +
  geom_area(colour = "white", linewidth = 0.2) +
  scale_fill_manual(values = pal, name = NULL) +
  scale_y_continuous(labels = scales::label_percent(scale = 1, accuracy = 1),
                     expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(1970, 2025, 5), expand = c(0, 0)) +
  labs(x = NULL, y = "Share of new-car market",
       title = "European new-car market by brand origin, 1970–2025") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank())

ggplotly(p_area, tooltip = "text")
```

### All five origins as trend lines

```{r lines}
p_line <- long %>%
  ggplot(aes(year, share, colour = origin,
             text = paste0(origin, ": ", round(share, 1), "%  (", year, ")"),
             group = origin)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = pal, name = NULL) +
  scale_y_continuous(labels = scales::label_percent(scale = 1, accuracy = 1)) +
  scale_x_continuous(breaks = seq(1970, 2025, 5)) +
  labs(x = NULL, y = "Share of new-car market",
       title = "European new-car market by brand origin, 1970–2025") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank())

ggplotly(p_line, tooltip = "text")
```

### The Asian challengers, close up

European brands dwarf the rest, so here are the three Asian origins on their own
scale.

```{r lines_asia}
p_line_asia <- long %>%
  filter(origin %in% c("Japanese", "Korean", "Chinese")) %>%
  ggplot(aes(year, share, colour = origin,
             text = paste0(origin, ": ", round(share, 1), "%  (", year, ")"),
             group = origin)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = pal, name = NULL) +
  scale_y_continuous(labels = scales::label_percent(scale = 1, accuracy = 1)) +
  scale_x_continuous(breaks = seq(1970, 2025, 5)) +
  labs(x = NULL, y = "Share of new-car market",
       title = "Japanese, Korean and Chinese brands in Europe") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank())

ggplotly(p_line_asia, tooltip = "text")
```

## What the curves show

**Japanese brands — the capped invasion.** Toyota, Nissan, Honda and Mazda climbed
fast in the early 1980s, then ran into a wall of bilateral quotas (Italy, France, the
UK) and the EU–Japan "Elements of Consensus" that governed access from 1991 to 1999.
Their answer was to build inside Europe. Share plateaued in the low teens, peaked
around **13%** in the mid-2000s, and has drifted gently down since.

**Korean brands — the steady climber.** Near zero in the early 1990s. Kia alone
logged fifteen straight years of share gains from 1995 to 2010, and Hyundai+Kia
reached roughly **8–9%** by the early 2020s — rising while most mainstream European
rivals fell.

**Chinese brands — the fastest entry yet.** Effectively zero before 2020, then a jump
from **3.1% (2024) to 6.1% (2025)** — about 812,000 registrations — led by MG/SAIC,
BYD, Chery (Omoda/Jaecoo), Leapmotor and Xpeng. They passed US brands in Q2 2025 and
overtook Korea in Q4 2025. On the country-of-manufacture basis, China-built cars
(including Western brands) reached ~7% and topped one million units in 2025.

**European brands — dominant but eroding.** Still the clear majority (VW Group,
Stellantis, Renault, BMW, Mercedes, Volvo), but their combined share has slipped
steadily as each Asian wave arrived.

**Japan then vs China now.** Extending the series to 1970 lets the two entries sit
side by side. Japanese brands went from roughly 1.5% in 1970 to about 9% by 1980 — a
decade-long climb that then hit political resistance and quotas and plateaued in the
low teens. Chinese brands have covered comparable ground far faster: near zero in
2020 to about 6% by 2025, half a decade rather than a full one. Whether they run into
the same ceiling — this time tariffs and local-content pressure rather than voluntary
export restraints — is the open question the chart sets up.

## Data table

```{r table}
df %>%
  filter(year %in% c(1970, 1980, 1990, 2000, 2010, 2020, 2024, 2025)) %>%
  transmute(Year = year,
            European = round(European, 1),
            Japanese = round(Japanese, 1),
            Korean = round(Korean, 1),
            Chinese = round(Chinese, 1),
            `US/Other` = round(`US/Other`, 1))
```

## Caveats

The series stitches together anchor points from several sources and interpolates
between them, so mid-decade values are estimates, not counts. "Europe" mixes
Western-Europe and later EU/EFTA+UK definitions across the period. And, again, the
**brand-nationality** shares here are not comparable to ACEA's much smaller
**import-origin** shares (Japan 4%, Korea 3%, China 7% in 2025) — those count where
cars are built, not whose badge they wear.

## Sources

- ACEA — *Economic and Market Report, Full Year 2025.*
  <https://www.acea.auto/publication/economic-and-market-report-global-and-eu-auto-industry-full-year-2025/>
- JATO Dynamics — *Beyond the brand: China's growing auto market share in Europe.*
  <https://www.jato.com/beyond-the-brand-chinas-growing-auto-market-share-in-europe>
- SCMP — *Chinese carmakers squeeze out Asian rivals in Europe.*
  <https://www.scmp.com/business/china-evs/article/3349794/chinese-carmakers-squeeze-out-asian-rivals-europe-exports-top-1-million-units>
- Harvard Business School — *Trade and the Single Car Market: the EC–Japan Elements of Consensus, 1985–1999.*
  <https://www.hbs.edu/ris/Publication%20Files/21-145_8a51f526-0033-41af-8f09-c449cc50bf24.pdf>
- Curbside Classic — *Why the Japanese invasion of Europe never happened.*
  <https://www.curbsideclassic.com/blog/news/curbside-newsstand-why-the-japanese-invasion-of-europe-never-happened/>
- CarBuzz — *The evolution of Kia and Hyundai.*
  <https://carbuzz.com/evolution-of-kia-and-hyundai/>
- Korea Herald — *Hyundai, Kia post record sales in Europe.*
  <https://www.koreaherald.com/view.php?ud=20240118000774>
